---
title: "Factor Investing with Alternative Data"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)

load("data_ml.RData")

data_ml %>%
  filter(date > "1999-12-31",
         date < "2019-01-01") %>%
  arrange(stock_id, date)
        
```
This dataset comprises information on 1,207 stocks listed in the US (possibly originating from Canada or Mexico). The time range starts in November 1998 and ends in March 2019. For each point in time, 93 characteristics describe the firms in the sample.

Our first observation is regarding the size factor, which states that companies with small capitalization tends grant higher returns than large capitalization. For that we create an equally weighted portofolio and summarize the mean return.
```{r}
data_ml %>% 
  group_by(date) %>% 
  mutate(large = Mkt_Cap_12M_Usd > median(Mkt_Cap_12M_Usd)) %>%  # Creates a Large Cap colum
  ungroup() %>% 
  mutate(year = lubridate::year(date)) %>%                       # Creates a year variable
  group_by(year, large) %>% 
  summarize(avg_return = mean(R1M_Usd)) %>%                      # Compute the avg return
  
  #Plot
  
  ggplot(aes(x= year, y= avg_return, fill = large)) +
  geom_col(position = "dodge") + theme_light() +         
  theme(legend.position = c(0.8, 0.2)) +
  coord_fixed(124) + theme(legend.title = element_blank()) +
  scale_fill_manual(values = c("#F87E1F", "#0570EA"), name = "",
                    labels = c("Small", "Large"))

```

```{r}
library(quantmod) # Data extraction
library(xtable)   # LaTex exports

min_date <- "1963-07-31"
max_date <- "2022-12-31"

temp <- tempfile()

kf_website <- "http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/"
kf_file <- "ftp/F-F_Research_Data_5_Factors_2x3_CSV.zip"

link <- paste0(kf_website, kf_file) # Linking the sites

download.file(link, temp, quiet = TRUE) #Download the site!

read_csv(unz(temp, "F-F_Research_Data_5_Factors_2x3.csv"), skip = 2) %>% 
  rename(date = `...1`, MKT_RF = `Mkt-RF`) %>% 
  mutate_at(vars(-date), as.numeric) %>%                                  # Convert valeus to numbers
  mutate(date = ymd(parse_date_time(date, "%Y%m"))) %>%                   # Right format
  mutate(date = rollback(date + months(1))) %>%                           # End of month date
  mutate( MKT_RF = MKT_RF / 100,
              SMB = SMB / 100,
              HML = HML / 100,
              RMW = RMW / 100,
              CMA = CMA / 100,
              RF = RF/100) %>% 
  filter(date >= min_date, date <= max_date) %>% 
  mutate(date = year(date)) %>%                                           # turn date into year
  gather(key = factor, value = value, - date) %>%                         # tidy shape
  group_by(date, factor) %>% 
  summarise(value = mean(value)) %>% 
  
  ggplot(aes(x=date, y=value, color = factor)) +
  geom_line()
 
 

```